---
title: "p8105_hw6_xz3312"
author: "Eric Zhang"
date: "2025-12-03"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(broom)
library(p8105.datasets)
library(modelr)
data("weather_df")
```

# Problem 1 
The Washington Post has gathered data on homicides in 50 large U.S. cities and made the data available through a GitHub repository

```{r}
# load data from GitHub

homicides_data <- read_csv(
  "https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"
)
```
Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. For this problem, limit your analysis those for whom victim_race is white or black. Be sure that victim_age is numeric.

```{r}
homicides_clean <- homicides_data %>% 
  mutate(
    city_state  = str_c(city, state, sep = ", "),
    resolved    = if_else(disposition == "Closed by arrest", 1, 0),

    # fix victim_age before numeric conversion
    victim_age  = na_if(victim_age, "Unknown"),
    victim_age  = as.numeric(victim_age),

    # convert sex and race to factors
    victim_sex  = factor(victim_sex),
    victim_race = factor(victim_race),

    # ensure female is the reference level
    victim_sex  = fct_relevel(victim_sex, "Female")
  ) %>% 
  filter(
    !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
    victim_race %in% c("White", "Black")
  ) %>% 
  drop_na(resolved, victim_age, victim_sex, victim_race)
```

For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.
```{r}
baltimore <- homicides_clean %>% 
  filter(city_state == "Baltimore, MD")

fit_baltimore <- glm(
  resolved ~ victim_age + victim_sex + victim_race,
  data = baltimore,
  family = binomial()
)

baltimore_or <- fit_baltimore %>% 
  broom::tidy(conf.int = TRUE, exponentiate = TRUE) %>% 
  filter(term == "victim_sexMale") %>% 
  select(term, estimate, conf.low, conf.high)

baltimore_or
```
Using the fitted logistic regression model for Baltimore, MD, the adjusted odds ratio for solving a homicide for male victims compared to female victims, holding age and race constant, is
`r round(baltimore_or$estimate, 2)`,
with a 95% confidence interval (`r round(baltimore_or$conf.low, 2)`, `r round(baltimore_or$conf.high, 2)`).

Since this odds ratio is below 1, the model suggests that homicides involving male victims are less likely to be solved than those involving female victims, after adjusting for age and race.

