---
title: "p8105_hw6_xz3312"
author: "Eric Zhang"
date: "2025-12-01"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(broom)
library(p8105.datasets)
library(modelr)
data("weather_df")
```

# Problem 1 
The Washington Post has gathered data on homicides in 50 large U.S. cities and made the data available through a GitHub repository

```{r}
# load data from GitHub

homicides_data <- read_csv(
  "https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"
)
```
### Part a
Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. For this problem, limit your analysis those for whom victim_race is white or black. Be sure that victim_age is numeric.

```{r}
homicides_clean <- homicides_data %>% 
  mutate(
    city_state  = str_c(city, state, sep = ", "),
    resolved    = if_else(disposition == "Closed by arrest", 1, 0),

    # fix victim_age before numeric conversion
    victim_age  = na_if(victim_age, "Unknown"),
    victim_age  = as.numeric(victim_age),

    # convert sex and race to factors
    victim_sex  = factor(victim_sex),
    victim_race = factor(victim_race),

    # ensure female is the reference level
    victim_sex  = fct_relevel(victim_sex, "Female")
  ) %>% 
  filter(
    !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
    victim_race %in% c("White", "Black")
  ) %>% 
  drop_na(resolved, victim_age, victim_sex, victim_race)
```

### Part b
For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors. Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.
```{r}
baltimore <- homicides_clean %>% 
  filter(city_state == "Baltimore, MD")

fit_baltimore <- glm(
  resolved ~ victim_age + victim_sex + victim_race,
  data = baltimore,
  family = binomial()
)

baltimore_or <- fit_baltimore %>% 
  broom::tidy(conf.int = TRUE, exponentiate = TRUE) %>% 
  filter(term == "victim_sexMale") %>% 
  select(term, estimate, conf.low, conf.high)

baltimore_or
```
Using the fitted logistic regression model for Baltimore, MD, the adjusted odds ratio for solving a homicide for male victims compared to female victims, holding age and race constant, is
`r round(baltimore_or$estimate, 2)`,
with a 95% confidence interval (`r round(baltimore_or$conf.low, 2)`, `r round(baltimore_or$conf.high, 2)`).

Since this odds ratio is below 1, the model suggests that homicides involving male victims are less likely to be solved than those involving female victims, after adjusting for age and race.

### Part c
Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims. Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.
```{r}
city_ors <- homicides_clean %>% 
  nest(data = -city_state) %>% 
  mutate(
    fit = map(
      data,
      ~ glm(
        resolved ~ victim_age + victim_sex + victim_race,
        data   = .x,
        family = binomial()
      )
    ),
    tidy = map(
      fit,
      ~ tidy(.x, conf.int = TRUE, exponentiate = TRUE)
    )
  ) %>% 
  select(city_state, tidy) %>% 
  unnest(tidy) %>% 
  filter(term == "victim_sexMale") %>% 
  select(city_state, estimate, conf.low, conf.high)

city_ors
```
Across cities, the adjusted odds ratios comparing male to female victims vary
substantially. For example, Albuquerque has an OR of `r round(city_ors$estimate[city_ors$city_state == "Albuquerque, NM"], 2)`
with a 95% CI of (`r round(city_ors$conf.low[city_ors$city_state == "Albuquerque, NM"], 2)`,
`r round(city_ors$conf.high[city_ors$city_state == "Albuquerque, NM"], 2)`), indicating higher odds of a homicide being solved
for male victims. In contrast, Baltimore shows a much lower OR of
`r round(city_ors$estimate[city_ors$city_state == "Baltimore, MD"], 2)` (95% CI:
`r round(city_ors$conf.low[city_ors$city_state == "Baltimore, MD"], 2)`,
`r round(city_ors$conf.high[city_ors$city_state == "Baltimore, MD"], 2)`),
suggesting male-victim homicides are less likely to be solved than female-victim homicides.

Many cities, such as Boston (`r round(city_ors$estimate[city_ors$city_state == "Boston, MA"], 2)`), have odds ratios below 1,
while others, such as Atlanta (`r round(city_ors$estimate[city_ors$city_state == "Atlanta, GA"], 2)`), are close to 1,
indicating little difference by victim sex. Overall, most confidence intervals include values below 1,
suggesting that in many cities male-victim homicides tend to be solved at lower rates.

### Part d
Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot.
```{r}
city_ors %>% 
  ggplot(aes(
    x = reorder(city_state, estimate),
    y = estimate,
    ymin = conf.low,
    ymax = conf.high
  )) +
  geom_pointrange() +
  geom_hline(yintercept = 1, linetype = "dashed") +
  coord_flip() +
  labs(
    x = "City",
    y = "Adjusted OR (Male vs Female)",
    title = "Odds of homicide being solved: male vs female victims"
  )
```
The plot shows substantial variation across cities in the odds of solving homicides for male versus female victims. Most cities have odds ratios below 1, suggesting that homicides involving male victims are generally less likely to be solved. Only a few cities have odds ratios above 1, and many confidence intervals are wide, indicating uncertainty in estimates of several cities.

# Problem 2
For this problem, we’ll use the Central Park weather data.

### Part a
The code chunk below will import these data from the p8105.datasets package.We’ll focus on a simple linear regression with tmax as the response with tmin and prcp as the predictors, and are interested in the distribution of two quantities estimated from these data:

r_hat_square
beta_hat_1 / beta_hat_2
```{r}
set.seed(1)

boot_results <- weather_df %>% 
  modelr::bootstrap(5000) %>% 
  mutate(
    # fit tmax ~ tmin + prcp in each bootstrap sample
    model  = map(strap, ~ lm(tmax ~ tmin + prcp, data = .x)),
    gl     = map(model, broom::glance),
    td     = map(model, broom::tidy),

    # r-hat squared from each model
    r_hat_sq = map_dbl(gl, "r.squared"),

    # β1-hat * β2-hat: coefficients for tmin and prcp
    beta_prod = map_dbl(
      td,
      ~ {
        b1 <- .x$estimate[.x$term == "tmin"]
        b2 <- .x$estimate[.x$term == "prcp"]
        b1 * b2
      }
    )
  ) %>% 
  select(r_hat_sq, beta_prod)

head(boot_results)
```
Across the 5000 bootstrap samples, the estimates of `r^2` remain very stable,
with a mean around `r round(mean(boot_results$r_hat_sq), 3)` and most values
close to `r round(boot_results$r_hat_sq[1], 3)`, indicating that the model
explains a large proportion of the variability in *tmax*. In contrast, the
product `β̂1 β̂2` is consistently small and negative, with typical values near
`r round(boot_results$beta_prod[1], 5)`, suggesting a weak combined effect of
*tmin* and *prcp* on *tmax* across bootstrap samples.

### Part b



